The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2009-07-29 script/de/chapters/numbertheory.tex
CT script, en: some minor changes. And moved sage sample code in the primes chapter from the appendix to a normal section (de still todo)

\begin{tabular}{|c|c|l|}
\hline
i  & $11^i{\rm ~mod~}347$ & \\
\hline
      0  &          1   &  \\
      1  &         11   &  \\                                     
      2  &        121   &  \\                                     
      3  &        290   &  \\                                     
     67  &         49   & gesuchter Exponent  \\                    
    172  &        284   &  \\                                                  
    173  &          1   &= Multiplikative Ordnung von $11 {\rm  ~(mod~} 347)$ \\ 
    174  &         11   &  \\                                                     
    175  &        121   &  \\                                     
    176  &        290   &  \\                                     
    240  &         49   & gesuchter Exponent  \\                     
\hline
\end{tabular}
\vskip +6 pt

>>>>

\begin{tabular}{|c|c|l|}
\hline
i  & $11^i \bmod 347$ & \\
\hline
      0  &          1   &  \\
      1  &         11   &  \\                                     
      2  &        121   &  \\                                     
      3  &        290   &  \\                                     
     67  &         49   & gesuchter Exponent  \\                    
    172  &        284   &  \\                                                  
    173  &          1   &= Multiplikative Ordnung von $11^i ~(\bmod 347)$  \\ 
    174  &         11   &  \\                                                     
    175  &        121   &  \\                                     
    176  &        290   &  \\                                     
    240  &         49   & gesuchter Exponent  \\                     
\hline
\end{tabular}
\vskip +6 pt

### 2009-07-29 script/de/chapters/primes.tex
CT script, en: some minor changes. And moved sage sample code in the primes chapter from the appendix to a normal section (de still todo)

\begin{sagecode}
\begin{Verbatim}%
[fontsize=\footnotesize,fontshape=tt]
E = [10, 40, 56, 64, 80, 90, 112, 128, 150, 160, 192, 256, 1024, 2048]
for e in E:
    # print "2^" + str(e), "---", 1.0*(2^e)  
    # print "2^%4d --- %0.6g" % (e, 2^e )   # nicer layout, but overflow at "2^1024 --- inf"
    print "2^%4d" % e , " --- ", 1.0*(2^e)

2^  10  ---  1024.00000000000
2^  40  ---  1.09951162777600e12
2^  56  ---  7.20575940379279e16
2^  64  ---  1.84467440737096e19
2^  80  ---  1.20892581961463e24
2^  90  ---  1.23794003928538e27
2^ 112  ---  5.19229685853483e33
2^ 128  ---  3.40282366920938e38
2^ 150  ---  1.42724769270596e45
2^ 160  ---  1.46150163733090e48
2^ 192  ---  6.27710173538668e57
2^ 256  ---  1.15792089237316e77
2^1024  ---  1.79769313486232e308
2^2048  ---  3.23170060713110e616
\end{Verbatim}
\caption{Spezielle Werte des Zweier- und Zehnersystems}
\end{sagecode}

>>>>

\begin{sagecode}
\begin{Verbatim}%
[fontsize=\footnotesize,fontshape=tt]
E = [10, 40, 56, 64, 80, 90, 112, 128, 150, 160, 192, 256, 1024, 2048]
for e in E:
    # print "2^" + str(e), "---", 1.0*(2^e)
    print "2^%4d" % e , " --- ", RR(2^e).n(24)
....:
2^  10  ---  1024.00
2^  40  ---  1.09951e12
2^  56  ---  7.20576e16
2^  64  ---  1.84467e19
2^  80  ---  1.20893e24
2^  90  ---  1.23794e27
2^ 112  ---  5.19230e33
2^ 128  ---  3.40282e38
2^ 150  ---  1.42725e45
2^ 160  ---  1.46150e48
2^ 192  ---  6.27710e57
2^ 256  ---  1.15792e77
2^1024  ---  1.79769e308
2^2048  ---  3.23170e616
\end{Verbatim}
\caption{Spezielle Werte des Zweier- und Zehnersystems}
\end{sagecode}

### 2009-07-29 script/en/chapters/numbertheory.tex
CT script, en: some minor changes. And moved sage sample code in the primes chapter from the appendix to a normal section (de still todo)

\begin{tabular}{|c|c|l|}
\hline
i  & $11^i \bmod 347$ & \\
\hline
      0  &          1   &  \\
      1  &         11   &  \\                                     
      2  &        121   &  \\                                     
      3  &        290   &  \\                                     
     67  &         49   & searched exponent \\                    
    172  &        284   &  \\                                                  
    173  &          1   &= multiplicative order of $11 \bmod 347$ \\ 
    174  &         11   &  \\                                                     
    175  &        121   &  \\                                     
    176  &        290   &  \\                                     
    240  &         49   & searched exponent \\
\hline
\end{tabular}
\vskip +6 pt

>>>>

\begin{tabular}{|c|c|l|}
\hline
i  & $11^i \bmod 347$ & \\
\hline
      0  &          1   &  \\
      1  &         11   &  \\                                     
      2  &        121   &  \\                                     
      3  &        290   &  \\                                     
     67  &         49   & searched exponent \\                    
    172  &        284   &  \\                                                  
    173  &          1   &= multiplicative order of $11^i \bmod 347$ \\ 
    174  &         11   &  \\                                                     
    175  &        121   &  \\                                     
    176  &        290   &  \\                                     
    240  &         49   & searched exponent \\
\hline
\end{tabular}
\vskip +6 pt

### 2009-07-29 script/en/chapters/primes.tex
added argument to example and remark definitions
CT script, en: some minor changes. And moved sage sample code in the primes chapter from the appendix to a normal section (de still todo)

\vskip +10pt
\begin{remark} With gaps, extremely large prime numbers were discovered
at an early stage.  \\
\end{remark}

@@

\begin{sagecode}
\begin{Verbatim}%
[fontsize=\footnotesize,fontshape=tt]
E = [10, 40, 56, 64, 80, 90, 112, 128, 150, 160, 192, 256, 1024, 2048]
for e in E:
    # print "2^" + str(e), "---", 1.0*(2^e)  
    # print "2^%4d --- %0.6g" % (e, 2^e )   # nicer layout, but overflow at "2^1024 --- inf"
    print "2^%4d" % e , " --- ", 1.0*(2^e)

2^  10  ---  1024.00000000000
2^  40  ---  1.09951162777600e12
2^  56  ---  7.20575940379279e16
2^  64  ---  1.84467440737096e19
2^  80  ---  1.20892581961463e24
2^  90  ---  1.23794003928538e27
2^ 112  ---  5.19229685853483e33
2^ 128  ---  3.40282366920938e38
2^ 150  ---  1.42724769270596e45
2^ 160  ---  1.46150163733090e48
2^ 192  ---  6.27710173538668e57
2^ 256  ---  1.15792089237316e77
2^1024  ---  1.79769313486232e308
2^2048  ---  3.23170060713110e616
\end{Verbatim}
\caption{Special values of the binary and decimal systems}
\end{sagecode}

@@

% --------------------------------------------------------------------------
\newpage
\hypertarget{Primes_Appendix_A}{}
\section*{Appendix A: Examples using Sage}
\addcontentsline{toc}{section}{Appendix A: Examples using Sage}
\index{Sage}

This appendix contains Sage source code to verify primality of
integers generated by the function $f(n) = n^2 - 9n + 61$. The code
defines a function called \verb!quadratic_prime_formula()! that takes
three arguments:
\begin{itemize}
\item \verb!start! --- An integer which is the lower bound for
  integers in the sequence $\texttt{start}, \texttt{start} + 1,
  \texttt{start} + 2, \dots, \texttt{end} - 1, \texttt{end}$.

>>>>

\vskip +10pt
\begin{remark}{:}\\
With gaps, extremely large prime numbers were discovered
at an early stage.
\end{remark}

@@

\begin{sagecode}
\begin{Verbatim}%
[fontsize=\footnotesize,fontshape=tt]
E = [10, 40, 56, 64, 80, 90, 112, 128, 150, 160, 192, 256, 1024, 2048]
for e in E:
    # print "2^" + str(e), "---", 1.0*(2^e)
    print "2^%4d" % e , " --- ", RR(2^e).n(24)
....:
2^  10  ---  1024.00
2^  40  ---  1.09951e12
2^  56  ---  7.20576e16
2^  64  ---  1.84467e19
2^  80  ---  1.20893e24
2^  90  ---  1.23794e27
2^ 112  ---  5.19230e33
2^ 128  ---  3.40282e38
2^ 150  ---  1.42725e45
2^ 160  ---  1.46150e48
2^ 192  ---  6.27710e57
2^ 256  ---  1.15792e77
2^1024  ---  1.79769e308
2^2048  ---  3.23170e616
\end{Verbatim}
\caption{Special values of the binary and decimal systems}
\end{sagecode}

% ---------------------------------------------------------------------------
% ---------------------------------------------------------------------------
\newpage
\hypertarget{Primes_Appendix_A}{}
\section{Examples using Sage}
% \section*{Appendix A: Examples using Sage}
% \addcontentsline{toc}{section}{Appendix A: Examples using Sage}
\label{Primes_Sage_samples}
\index{Sage!Code examples}
\index{Sage}

The following section contains Sage source code related to contents of 
the ``\nameref{Label_Kapitel_2}'' chapter (chap. no. \ref{Label_Kapitel_2}).

% ---------------------------------------------------------------------------
% \newpage
\subsection{Check primality of integers generated by quadratic functions}

The following Sage code verifies the primality of integers generated
by the function $f(n) = n^2 - 9n + 61$. The code defines a function called
\verb!quadratic_prime_formula()! that takes three arguments:
\begin{itemize}
\item \verb!start! --- An integer which is the lower bound for
  integers in the sequence $\texttt{start}, \texttt{start} + 1,
  \texttt{start} + 2, \dots, \texttt{end} - 1, \texttt{end}$.

@@
